---
title: "My first quarto"
author: "Jordan Vest"
date: today
format: 
  html:
    toc: true
---
## Introduction
This is my first *Quatro* document and I am going to plot some penguin data
You can convert markdown quarto to pdf, MS Word, presentation, git and many others
[HTML format](https://quarto.org/docs/reference/formats/html.html)
[YAML Codes](https://quarto.org/docs/reference/cells/cells-knitr.html)
Note that images, links, text lives outside chunk


```{r}
#| message: false #yaml tells it to hide the output on render

#Load libraries
library(tidyverse)
library(here)
library(palmerpenguins)
#cmd+enter will run the line of code you are on. green arrow will run the chunk

```
each chunk should only have 1 figure and tackle one task

## Look at the data
```{r}
glimpse(penguins)
```

## Make a plot
This is looking at the relationship between *bill length** and **bill depth**.

```{r}
penguins %>% #penguin dataset
  ggplot(aes(x=bill_length_mm, y = bill_depth_mm, color = species))+
  geom_point()
```
### Fun package
install.packages(beepr)
library(beepr)
beep(5)
```{r}

```


Workflow between obsidian and quarto? 
#Get better grades by...
Save your work then push
Clear Environment
Restart R environment to make sure you have the libraries loaded in script
.rs.restartR()
#remove whole list
rm(list = ls())

Watch that you don't rename plots and then try to save a plot that doesn't exist
Use the source to run through and see if there is error in script
Double check in github

```{r}

```

Look into jupiter/quarto lab notebook
#4parts to Quarto
Metadata
Text
Code
Output

#Metadata
---
key:value
title: "My awesome markdown file
format:
  html:
  toc: true #tbl of contents
  toc-depth: 4 #4 levels
description: A great new article
affiliation: RStudio
date: "'rSys.Date()'"
---
