callie: is this new knowledge enough?
---
subtitle: "Tidy Tuesday Week 13"
author: "Jordan Vest"
date: today
format: 
  html:
    toc: true
    toc-title: "Are you lost? Look here!" 
    theme: darkly
    smooth-scroll: true 
    respect-user-color-scheme: true 
    link-external-newwindow: true
    linkcolor: pink 
    page-layout: article
---
# New Knowledge
We talked about word clouds in class, but I didn’t make one during the exercises. So this week I learned to actually build one from scratch and used {tidytext} to tokenize the data and {wordcloud2} to visualize the results. 

## Load libraries
```{r}
#| message: false
#| warning: false
library(tidyverse)
library(tidytext)     
library(wordcloud2)
```

## Load data
```{r}
#| message: false
#| warning: false
holmes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-11-18/holmes.csv')
```

## Clean data

### Detect all lines that are in quotes 
This is a very rough way to filter out the data we want but since Watson narrates these stories, much of the dialogue in quotes will give us the most spoken words.
```{r}
#| message: false
#| warning: false
holmes_lines <- holmes %>%
  filter(!is.na(text)) %>%
  filter(str_detect(text, '"'))  # keeps lines with double quotes
```

### Tokenize words
```{r}
#| message: false
#| warning: false
holmes_words <- holmes_lines %>%
  unnest_tokens(word, text)
```

### Filter out stop words
```{r}
#| message: false
#| warning: false
holmes_words_clean <- holmes_words %>%
  filter(!word %in% tidytext::stop_words$word) %>%
  filter(str_detect(word, "^[a-z']+$"))
```

### Count most said words
```{r}
#| message: false
#| warning: false
holmes_word_counts <- holmes_words_clean %>%
  count(word, sort = TRUE)
```

### Top 100 words
```{r}
#| message: false
#| warning: false
top_n <- 100
holmes_top <- holmes_word_counts %>% slice_max(n = top_n, order_by = n)
```


# Make a word cloud
```{r}
#| message: false
#| warning: false
wc <- wordcloud2(data = holmes_top, size = 1, color = "random-dark", backgroundColor = "white")
wc
```

This is cool but I feel I didn't learn any flashy "new" thing so I'm going to add on to it by putting the worldcloud inside a sherlock silhouette

```{r}
```{r}
figPath <- ("/Users/jordanvest/Desktop/repositories/Vest/Tidy_Tuesday/Data/sherlock.png")
```

```{r}
wc <- wordcloud2(
  data = holmes_top, 
  size = 1, 
  color = "random-dark", 
  backgroundColor = "white",
  figPath = figPath
)
```