---
title: "Wk7_LectureNotes"
author: "Jordan Vest"
format: html
---
check out Ruski HW6 for inspo

# Intro to Maps
Other very helpful packages:

sf (This is for the serious GIS people. If you want to do actual spatial analyses this is the package to use)
ggmap (Uses the google maps API to make maps. Really easy to use, but requires a credit card for the API...)
tmap (static and interactive maps)
leaflet (interactive html maps-- Awesome to use with rmakrdown files)
Helpful references:

The required reading for today
[Advanced maps](https://r.geocompx.org/adv-map.html) 
[Bathymetry example](https://ben-williams.github.io/updated_ggplot_figures.html#maps)
[Earth Lab](https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/understand-epsg-wkt-and-other-crs-definition-file-types/)



# Install packages
```{r}
install.packages(c("maps", "mapproj", "mapdata", "here","tidyverse"))
```
# Load Libraries
```{r}
# Load libraries
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
```
# Read in Data
```{r}
# Read in data on population in California by county
popdata<-read_csv(here("Week_07","data","CApopdata.csv"))
#read in data on number of seastars at different field sites
stars<-read_csv(here("Week_07","data","stars.csv"))
```
# {maps} package
```{r}
# get data for the entire world
world<-map_data("world")
head(world)
```
```{r}
# get data for the USA
usa<-map_data("usa")
head(usa)
```
```{r}
# get data for states
states<-map_data("state")
head(states)
```

```{r}
# get data for counties
counties<-map_data("county")
head(counties)
```

```{r}

```

