---
title: "Wk7_HWa"
author: "Jordan Vest"
description: A great map of Bigfoot sightings across the USA
format: html
---
# Assignment
make a map using Quarto with one of the following datasets from past Tidy Tuesdays. Make sure that scripts and output are saved appropriately.

* 2 points - best Quarto practices

* 2 points - best file structure practices

* 2 points - best scripting practices

* 4 points - best map practices

### Load Libraries
```{r}
#| message: false
#| warning: false
library(tidytuesdayR)
library(ggplot2)
library(tidyverse)
library(maps)
library(mapdata)
library(mapproj)
library(here)
```

### Read in data
The data  comes from Bigfoot Field Researchers Organization (BFRO) by way of Data.World.

```{r}
#| message: false
#| warning: false
bigfoot_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-09-13/bigfoot.csv")

# Clean the data: remove rows missing key info like state or coordinates
bigfoot_clean <- bigfoot_raw %>%
  filter(!is.na(state), !is.na(latitude), !is.na(longitude))

# Save cleaned data to my Data folder
write_csv(bigfoot_clean, here("Week_07", "Data", "bigfoot_clean.csv"))

# Re-load from local file for reproducibility
bigfoot <- read_csv(here("Week_07", "Data", "bigfoot_clean.csv"))
```

### Wrangle Data
```{r}
#| message: false
#| warning: false
bigfoot_state <- bigfoot %>%
  mutate(state = tolower(state)) %>%     # make lowercase to match map_data
  group_by(state) %>%
  summarise(n_sightings = n())
```
#### Get map data
```{r}
#| message: false
#| warning: false
usa <- map_data("state")
```
#### Join map with sightings
```{r}
#| message: false
#| warning: false
bigfoot_map <- usa %>%
  left_join(bigfoot_state, by = c("region" = "state"))
```

### Make a map
```{r}
#| message: false
#| warning: false
ggplot() +
  geom_polygon(data = bigfoot_map, 
               aes(x = long, 
                   y = lat, 
                   group = group, 
                   fill = n_sightings),
               color = "white",
               linewidth = 0.3) +
  # Projection and limits for continental U.S.
  coord_map("albers", lat0 = 30, lat1 = 45) +
  # Log scale because sightings vary a lot by state
  scale_fill_viridis_c(
    option = "cividis",
    trans = "log10",
    name = "Sightings (log scale)"
  ) +
  # Add some flair and polish
  labs(
    title = "Reported Bigfoot Sightings Across the United States",
    subtitle = "Based on publicly available Bigfoot Field Researchers Organization (BFRO) reports",
    caption = "Data: TidyTuesday (2022-10-13) • Visualization by Jordan Vest"
  ) +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "lightblue"),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 10),
    plot.caption = element_text(size = 8, hjust = 1),
    legend.position = c(0.85, 0.25),
    legend.background = element_rect(fill = alpha("white", 0.7)),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```

### Save it
```{r}
ggsave(here("Week_07","Output","Bigfoots_HW7a.pdf"))
```

