---
title: "Wk6_HW6"
author: "Jordan Vest"
format: gfm
prefer-html: true
knitr:
  opts_chunk:
    fig.path: "../Output/"
---

# Introduction

Today we are practicing with Quarto and completing the following tasks.

- Data import 
- Summary Table 
- Plot 

## Load the libraries

```{r}
#| message: false
#| warning: false

#Load libraries
library(tidyverse)
library(here)
library(kableExtra)

# Read in the data
here::i_am("Week_06/Script/Wk6_HW6_Script_final.qmd")
chemicaldata <- read_csv(here("Week_06", "Data", "chemicaldata_maunalua copy.csv"))

# Peek at the first few rows
head(chemicaldata)
```

# Make a summary table
```{r}
#| warning: false
#| message: false
#| label: tbl-summary
#| tbl-cap: "Data Summary Table"

summary_table <- chemicaldata %>%
  group_by(Zone) %>%
  summarise(
    mean_temp = mean(Temp_in, na.rm = TRUE),
    mean_sal = mean(Salinity, na.rm = TRUE),
    n = n()
  )

summary_table %>%
  kbl(caption = "Average Temperature and Salinity by Zone", digits = 2) %>%
  kable_styling(full_width = FALSE, position = "center")
  

```
@tbl-summary gives average _*Temperature*_ and _*Salinity*_ values for the _*Zones*_ that have data

# Make a plot
```{r}
#| warning: false
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4
#| dev: "png"
#| label: fig-scatter
#| fig-cap: "Temperature v Salinity plot"

print(
ggplot(chemicaldata, aes(x = Salinity, y = Temp_in, color = Zone)) +
  geom_point(size = 3, alpha = 0.7) +
  coord_cartesian(xlim = c(25, 35)) + 
  theme_minimal(base_size = 12) +
  labs(
    title = "Temperature vs. Salinity by Zone",
    x = "Salinity (psu)",
    y = "Temperature (°C)",
    color = "Zone"
  ))

```

When you look @fig-scatter we can see there is a distribution of _*Salinity*_ values in the *Diffuse* and *Transition* zones.